---
title: "Dissertation 2018-19"
author: "Jack Unsworth"
date: "18 October 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = T)
```

We use data from http://www.football-data.co.uk/englandm.php, this is updated weekly
The descriptions of variables are in the file http://www.football-data.co.uk/notes.txt

Loading packages required for the analysis:

```{r, include = FALSE}
# install.packages("devtools")
library(devtools)
# install.packages('ggplot2')
library(ggplot2)
```

Set the working directory, load in the data and combine it into one data frame:

```{r}
PL1516 <- read.csv('PL1516.csv')  # Loading Data
PL1617 <- read.csv('PL1617.csv')
PL1718 <- read.csv('PL1718.csv')
PL1819 <- read.csv('PL1819.csv')
PL <- rbind(PL1516,PL1617,PL1718,PL1819)
```

Data Analysis

We first look at the distribution of home and away goals, plot these and run a t test for differences:

```{r}

mean(PL$FTHG)  # We see the home and away goals are different values by a significant amount
mean(PL$FTAG)

par(mfrow = c(2,1))  # Changes number of graphs on plot
barplot(prop.table(table(PL$FTHG)), main = 'Home Goals', xlab = 'Goals', ylab = 'Frequency', ylim = c(0,0.4))
barplot(prop.table(table(PL$FTAG)), main = 'Away Goals', xlab = 'Goals', ylab = 'Frequency', ylim = c(0,0.4))
par(mfrow = c(1,1))

t.test(PL$FTHG,PL$FTAG,var.equal = FALSE)  # This gives statistical evidence of home advantage
```


Independent Poisson Model:

```{r, Include = T}
poisson_model <- 
  rbind(
    data.frame(goals=PL$FTHG,
               team=PL$HomeTeam,
               opponent=PL$AwayTeam,
               home=1),
    data.frame(goals=PL$FTAG,
               team=PL$AwayTeam,
               opponent=PL$HomeTeam,
               home=0)) 
  
Mod1 <- glm(goals ~ home + team + opponent, family=poisson(link=log), data=poisson_model)
```


There are two wrapper functions in seperate files. One takes in a model and fixtures and returns the predicted results, this is called 'Predict1'. The second takes in a model and returns the premier league table according to the model prediction for all matches int he season up to the current date, excluding week 1 which currently must be included within the training data to avoid new levels in the model causing issues.This second function is called 'LeagueTablePredict'.

The two functions are demonstrated below using the poisson linear model and in the first, data from game week 8 of the Premier League 2018/19.

```{r, include = T}
source("Predict1.R")
source("LeagueTablePredict.R") 

FixturesGW8 <- data.frame( Team = c('Brighton','Crystal Palace','Leicester','Man United','Southampton','Burnley','Watford','Tottenham','Fulham','Liverpool'), Opponent = c('West Ham','Wolves','Everton','Newcastle','Chelsea','Huddersfield','Bournemouth','Cardiff','Arsenal','Man City') )  # Home teams go first

FixturesGW9 <- data.frame( Team = c('Chelsea','West Ham','Newcastle','Bournemouth','Cardiff','Wolves','Man City','Huddersfield','Everton','Arsenal'), Opponent = c('Man United','Tottenham','Brighton','Southampton','Fulham','Watford','Burnley','Liverpool','Crystal Palace', 'Leicester') )

Predict1(FixturesGW9,Mod1)
LeagueTablePredict(Mod1)

```